Molecular transport in human pial perivasculature

### 2024-11-18 paper/supplementary_main.tex
Reached the numerics - woohoo!

\subsection{Mathematical models and definitions}
\label{sec:more_maths}

@@

The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
space averaged over each perivascular cross-section and is governed by 
\begin{equation}
  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s (\hat c) ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
\label{eq:1d_pde}
\end{equation}
where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
associated PVS $\Omega^i$, the area of the cross-section
$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
and $\hat{u}$ is a convective velocity representing an (average) CSF
flow velocity in the axial direction of the perivascular spaces. To
complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
boundary conditions at the end nodes. To this end, define the set of
internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
that are connected to two or more edges and the set of end nodes as
$\mathcal{T}$, write $\hat{c}^i = \hat{c}|_{\Lambda^i}$, and let
$\mathcal{E}(y)$ denote the set of (one, two or three) edges in
$\{\Lambda^i\}$ sharing the node $y$. To impose continuity of the
concentrations, we set that the concentrations when viewed from each
edge must match at nodes:
\begin{equation}
  \hat{c}^i (y) = \hat{c}_y \qquad \foralls \Lambda^i \in \mathcal{E}(y),
  \quad \foralls y \in \mathcal{B} .
\end{equation}
Moreover, to ensure conservation of flux, we set that the flux going
in and out at nodes should add to zero, or more precisely, that for
each $y \in \mathcal{B}$:
\begin{equation}
  \sum_{i \in \mathcal{E}(y)}
  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c^i}(y) \right ) n_i(y)
  = 0
  \label{eq:flux_conservation}
\end{equation}
where the function $n_i : \mathcal{V} \rightarrow \{-1, 0, 1\}$
defines an orientation of each vertex relative to the
edges. Specifically, for an edge $\Lambda^i$ from $y_{\rm in}$ to $y_{\rm out}$
\begin{equation}
  n_i(y_{\rm in}) = 1, \quad
  n_i(y_{\rm{out}}) = -1, \quad
  n_i(y) = 0 \quad \foralls y \in 
  \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} .  
\end{equation}

@@

The time-independent, incompressible Stokes equations model the flow
of an incompressible Newtonian fluid at low Reynolds numbers, and read
as follows: over the domain $\Omega_{\rm CSF} \subset \R^3$, find the
velocity vector field $\bm u : \Omega_{\rm CSF} \rightarrow \R^3$ and
the pressure $p : \Omega_{\rm CSF} \rightarrow \R$ such that
\begin{subequations}
  \begin{alignat}{2}
    - \mu \Delta \bm u  + \nabla p &=  0 \quad && \text{ in } \Omega_{\rm CSF},
    \label{eq:momnetum_equation} \\ 
    \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF},
    \label{eq:divergence_equation} .
  \end{alignat}
\end{subequations}
In addition, we impose the following boundary conditions to model flow
induced by CSF production in the choroid plexus with $\Gamma_{\rm AM-U}$ as the main efflux site:
\begin{subequations}
  \begin{alignat}{2}
    \mu \nabla \bm u \cdot \bm{n} - p \bm n &= -R_0 ( \bm u \cdot \bm n ) \bm n
    && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
    \bm u &= 0 && \text{ on } \Gamma_{\mathrm{AM-L}}, \\
    \bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
    \bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm \tau = 0 \quad && \text{ on } \Gamma_{\rm{LV}} ,  
  \end{alignat}
\end{subequations}
where $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$. We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:supp:dispersion}.

@@

To model flow induced by pressure differences between end nodes in a
perivascular network $\Lambda$, with segments $\Lambda^i$, internal
nodes $\mathcal{B}$ and end nodes $\mathcal{T}$, we consider the
following system of hydraulic network
equations~\cite{daversin2022geometrically, gjerde2024directional}. The
unknowns are the PVS velocity $u_{\rm pvs}$ and pressure $p_{\rm
  pvs}$, which represent the fluid velocity in the axial direction
averaged across the PVS cross-sections and the cross-section-average
fluid pressure, respectively, solving
\mer{MER: Stopped here.}
\begin{alignat}{2}
A \,  u_{\rm pvs}   + \frac{\kappa}{\mu} \, A \, \partial_{s} p_{\rm pvs} & = 0, &&  \quad \rm in  \,\, \Lambda  \\ 
-\partial_s (A \, u_{\rm pvs}) & = f, && \quad \rm in  \,\, \Lambda .  
\end{alignat} 
In the above, $A$ is the area of the PVS, $\mu$ is the dynamic viscosity, and $\kappa$ is derived from the assumption of Poiseuille
flow in the annular cross-section of the PVS \cite{daversin2022geometrically,tithof2022network}: 
\begin{equation}
\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
\end{equation}
On bifurcations, conservation of fluxes is enforced weakly and continuity of the pressure is enforced strongly (in the choice of the finite element spaces).

Since the pressure gradient the caused by the steady production-driven CSF flow (see \cref{sec:csf_fluid_vel}), we impose the precomputed fluid pressure $p_{\rm CSF}$ on the networks root nodes:

\begin{equation}
    p_{\rm PVS} = p_{\rm CSF} \quad \text{ on } \Lambda^{\rm out}
\end{equation}

In case that the root does not lie within the CSF-space but the parenchyma, the CSF pressure is not yet computed. For that reason, we apply a simple extension technique to map the CSF pressure field into $\Omega_{\rm PAR}$. In particular, we solve a Laplace equation on $\Omega_{\rm PAR}$ with the CSF pressure on the Pial membrane ($\Gamma_{\rm Pia}$) as a boundary condition.

\subsection{Numerical Method}
\label{sec:details_numerical_method}
Here we provide details on the numerical method used to solve the coupled system. We are given a mesh $\mathcal{T}_h $, consisting of elements denoted by $E$, of $\Omega= \Omega_{\rm PAR} \cup \Omega_{\rm CSF}$.  The collection of all interior facets to $\Omega_{\rm PAR}$ (resp.  to $\Omega_{\mathrm{CSF}}$) is  denoted by $\mathcal{F}_{h,b}$ (resp. $\mathcal{F}_{h,c}$).  The union of all interior facets $\mathcal{F}_i = \mathcal{F}_{h,b} \cup \mathcal{F}_{h,c}$. Facets lying on the pial interface are denoted by $\mathcal{F}_{\Gamma_{\rm pia}}$, on the skull by $\mathcal{F}_{\Gamma_{\rm skull}}$, and on the spinal canal by $\mathcal{F}_{\Gamma_{\rm SAS}}$. Note that $\mathcal{F}_i \cap \mathcal{F}_{\Gamma_{\rm pia}} = \emptyset $. The union of all the facets is denoted by $\mathcal{F}_h$. The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is denoted by $\mathcal{E}_{h}$.

@@

We follow \cite{hong2016robust} and approximate the velocity field $u$ and the pressure $p$ with the following finite element spaces. 
\begin{align}
\bm V_h  & = \{ \bm v  \in H(\mathrm{div}; \Omega_{\mathrm{CSF}}): \; \bm v \vert_{E} \in \mathbb{BDM}^2 (E),  \; E \in \mathcal{E}_h; \;  \bm v \cdot \bm n = 0 \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull^l}} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} \text{ on } \Gamma_{\mathrm{LV}} \} \\ 
S_h  & = \{q \in L^2(\Omega):  \;  q \vert_E \in \mathbb{P}^1(E),  \; E \in \mathcal{E}_h \}. 
\end{align}
Here, $H(\mathrm{div};\Omega)$ is the space of $L^2$ vector fields with $L^2$ divergence,  $\mathbb{BDM}^2$ is the Brezzi-Douglas--Marini element \cite{brezzi1987mixed} of order 2, and  $\bm n$ (resp. $\bm \tau$) is the unit outward normal (resp. tangent) vector.  
Given any vector $\bm v$, the normal and tangential components on each facet  are denoted by 
\[ 
\bm v_n = (\bm v \cdot \bm n)\bm n , \quad \bm v_t = (\bm v - \bm v_n). 
\]
Since $V_h \subset H(\mathrm{div}, \Omega_{\rm{CSF}})$, then $[\bm v_n] = 0 $ on $\mathcal{F}_{h,c}$. Continuity in the tangential component is enforced weakly via interior penalization. For notational simplicity, denote by $\mathcal{F}_{\Gamma}$ the facets lying on $\Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull}} \cup \Gamma_{\rm{SSAS}} \cup \Gamma_{\mathrm{LV}}$. Define the form 
\begin{multline}
\mathcal{A}_h (\bm u, \bm v) = \sum_{E \in \mathcal{E}_h} \int_{E} \mu \nabla u : \nabla v - \sum_{
F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma} 
} \int_{F} \mu \{\nabla u\}\cdot \bm n_F \cdot [\bm v_t] 
\\ - \sum_{
F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma} 
} \int_{F} \mu \{\nabla v\}\cdot \bm n_F \cdot [\bm u_t]
+ \sum_{
F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma} 
} \frac{\sigma \mu}{h} \int_{F} [\bm u_t ] \cdot [\bm v_t].   
\end{multline}
We set $\sigma =20$. The discretization for the CSF flow, is then to find $(\bm u_h, p_h) \in \bm V_h \times S_h$ such  that 
\begin{align}
\mathcal{A}_h(u_h, v_h)  + \sum_{F \in \mathcal{F}_{\Gamma_{\mathrm{skull}}}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n) - \int_{\Omega} \nabla \cdot \bm v_h \, p_h  & = 0, \quad \forall \bm v_h \in \bm V_h   \\ 
(\nabla \cdot \bm u_h , q_h ) & = 0, \quad \forall q_h \in S_h.
\end{align}
\subsubsection{PVS flow}

Recall that we compute pressure-driven flow within the PVS by solving a 1D Darcy equations in the PVS network. The unknowns are the averaged PVS velocity, $u_{\rm pvs} = \langle u_{v,s} \rangle$, and the cross-section pressure $p_{\rm pvs} $ (assumed to be constant in each cross-section). The system is given by  \cite{daversin2022geometrically, 
gjerde2024directional}

\begin{alignat}{2}
A \,  u_{\rm pvs}   + \frac{\kappa}{\mu} \, A \, \partial_{s} p_{\rm pvs} & = 0, &&  \quad \rm in  \,\, \Lambda  \\ 
-\partial_s (A \, u_{\rm pvs}) & = f, && \quad \rm in  \,\, \Lambda .  
\end{alignat}

where $A$ is the area of the PVS, $\mu$ is the dynamic viscosity, and $\kappa$ is derived from the assumption of Poiseuille flow in the annular cross-section of the PVS  \cite{daversin2022geometrically,tithof2022network}: 
\begin{equation}
\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
\end{equation}

As boundary conditions, we impose the precomputed fluid pressure $p_{\rm CSF}$ on the networks root nodes:
\begin{equation}
    p_{\rm PVS} = p_{\rm CSF} \quad \text{ on } \Lambda^{\rm out}
\end{equation}

In case that the root does not lie within the CSF-space but the parenchyma, the CSF pressure is not yet computed. For that reason, we apply a simple extension technique to map the CSF pressure field into $\Omega_{\rm PAR}$.
In particular, we solve the Laplace equation
\begin{equation}
    - \Delta p = 0; \quad p = p_{\rm CSF} \text{ on } \Gamma_{\rm Pia}
\end{equation}
to obtain the extension of $p_{\rm CSF}$ onto $\Omega_{\rm PAR}$.

@@

The weak form reads: Find $(u_{\rm PVS}, p_{\rm PVS}) \in V_h^{\rm PVS} \times Q_h^{\rm PVS}$ such that

\begin{align}
(A u_{\rm PVS}, v) + \left(v, \frac{A \kappa}{\mu} \partial_s p_{\rm PVS}  \right) &= 0 \quad \forall v \in V_h^{\rm PVS} \\
(A u_{\rm PVS} , \partial_s q ) &= 0 \quad  \forall q \in  Q_h^{\rm PVS}
\end{align}

@@

\subsection{Dispersion in the CSF spaces}
\label{sec:supp:dispersion}

>>>>

\subsection{Mixed-dimensional transport and flow equations}
\label{sec:more_maths}

@@

The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
space averaged over each perivascular cross-section and is governed by 
\begin{equation}
  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s (\hat c) ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
\label{eq:1d_pde}
\end{equation}
where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
associated PVS $\Omega^i$, the area of the cross-section
$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
and $\hat{u}$ is a convective velocity representing an (average) CSF
flow velocity in the axial direction of the perivascular spaces. To
complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
boundary conditions at the end nodes. To this end, define the set of
internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
that are connected to two or more edges and the set of end nodes as
$\mathcal{T}$, write $\hat{c}^i = \hat{c}|_{\Lambda^i}$, and let
$\mathcal{E}(y)$ denote the set of (one, two or three) edges in
$\{\Lambda^i\}$ sharing the node $y$. To impose continuity of the
concentrations, we set that the concentrations when viewed from each
edge must match at nodes:
\begin{equation}
  \hat{c}^i (y) = \hat{c}^j(y)
  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
  \qquad \foralls y \in \mathcal{B} .
\end{equation}
Moreover, to ensure \discuss{mass conservation}, we set that the flux
going in and out at nodes should add to zero, or more precisely, that
for each $y \in \mathcal{B}$:
\begin{equation}
  \sum_{\Lambda^i \in \mathcal{E}(y)}
  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n_i(y)
  = 0
  \label{eq:flux_conservation}
\end{equation}
where the function $n_i : \mathcal{V} \rightarrow \{-1, 0, 1\}$
defines an orientation of each vertex relative to the
edges. Specifically, for an edge $\Lambda^i$ from $y_{\rm in}$ to $y_{\rm out}$
\begin{equation}
  n_i(y_{\rm in}) = 1, \quad
  n_i(y_{\rm{out}}) = -1, \quad
  n_i(y) = 0 \quad \foralls y \in 
  \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} .  
\end{equation}

@@

The time-independent, incompressible Stokes equations model the flow
of an incompressible Newtonian fluid at low Reynolds numbers, and read
as follows: over the domain $\Omega_{\rm CSF} \subset \R^3$, find the
velocity vector field $\bm u : \Omega_{\rm CSF} \rightarrow \R^3$ and
the pressure $p : \Omega_{\rm CSF} \rightarrow \R$ such that
\begin{subequations}
  \begin{alignat}{2}
    - \mu \Delta \bm u  + \nabla p &=  0 \quad && \text{ in } \Omega_{\rm CSF},
    \label{eq:momnetum_equation} \\ 
    \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF},
    \label{eq:divergence_equation} .
  \end{alignat}
\end{subequations}
In addition, we impose the following boundary conditions to model flow
induced by CSF production in the choroid plexus with $\Gamma_{\rm AM-U}$ as the main efflux site:
\begin{subequations}
  \begin{alignat}{2}
    \mu \nabla \bm u \cdot \bm{n} - p \bm n &= -R_0 ( \bm u \cdot \bm n ) \bm n
    && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
    \bm u &= 0 && \text{ on } \Gamma_{\mathrm{AM-L}}, \\
    \bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
    \bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm \tau = 0 \quad && \text{ on } \Gamma_{\rm{LV}} ,  
  \end{alignat}
\end{subequations}
where $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the (dynamic) CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$. We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:supp:dispersion}.

@@

To model steady flow induced by pressure differences between end nodes
in a perivascular network $\Lambda$, with edges $\{\Lambda^i\}_i$,
internal nodes $\mathcal{B}$ and end nodes $\mathcal{T}$, we consider
the following system of hydraulic network
equations~\cite{tithof2022network, daversin2022geometrically,
  gjerde2024directional}. The unknowns are the PVS flux $\hat{q} :
\Lambda \rightarrow \R$ and pressure $\hat{p} : \Lambda
\rightarrow \R$, which represent the fluid flux and the average
pressure across cross-sections of the PVSs, respectively. These are
defined for each $\Lambda^i$ by solving:
\begin{subequations}
\begin{alignat}{2}
  \hat{q} + \frac{A \kappa}{\mu} \partial_{s} \hat{p} & = 0 &&
  \quad \text{ in } \Lambda^i , \\ 
  - \partial_s \hat{q} & = 0 &&
  \quad \text{ in } \Lambda^i ,  
\end{alignat} 
\label{eq:1d_flow}%
\end{subequations}%
where $A$ is the area of the PVS cross-sections, $\mu$ is the CSF viscosity as before, and $\kappa$ is derived from an assumption of Poiseuille flow in the annular cross-section of the PVS as~\cite{tithof2022network}: 
\begin{equation}
\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
\end{equation}
The PVS flux $\hat{q}$ and the associated average PVS velocity in the
axial direction $\hat{u}$ are directly related by 
area
\begin{equation}
  \hat{q}^i = A^i \hat{u}^i \qquad \foralls \Lambda^i.
\end{equation}

In addition, to complete~\eqref{eq:1d_flow}, we impose continuity of
the fluid pressure and conservation of flux at each internal node. Write
$\hat{p}^i = \hat{p}|_{\Lambda^i}$. These conditions then read as: for
each $y \in \mathcal{B}$ with connected edges $\mathcal{E}(y)$:
\begin{subequations}
\begin{align}
  \hat{p}^i(y) &= \hat{p}^j(y) \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y), \\
  \sum_{\Lambda^i \in \mathcal{E}(y)} \hat{q}^i n^i(y) &= 0  .
\end{align}
\end{subequations}
Finally, to drive this flow, we impose given fluid pressures $p_0$ at the
network end nodes:
\begin{equation}
  \hat{p}(y) = p_0(y) \qquad y \in \mathcal{T} .
\end{equation}
In the case where modelling flow in the perivascular networks
$\Lambda_a, \Lambda_v$ induced by CSF production, we impose the fluid
pressure $p_{\rm CSF}$ computed in the CSF spaces $\Omega_{\rm CSF}$
(and its harmonic extensions into $\Omega_{\rm PAR}$) as the given
fluid pressures $p_0$.

\mer{MER: Stopped here.}

\subsection{Numerical solution of the transport and flow equations}
\label{sec:details_numerical_method}

Here we provide details on the numerical method used to solve the coupled system. We are given a mesh $\mathcal{T}_h $, consisting of elements denoted by $E$, of $\Omega= \Omega_{\rm PAR} \cup \Omega_{\rm CSF}$.  The collection of all interior facets to $\Omega_{\rm PAR}$ (resp.  to $\Omega_{\mathrm{CSF}}$) is  denoted by $\mathcal{F}_{h,b}$ (resp. $\mathcal{F}_{h,c}$).  The union of all interior facets $\mathcal{F}_i = \mathcal{F}_{h,b} \cup \mathcal{F}_{h,c}$. Facets lying on the pial interface are denoted by $\mathcal{F}_{\Gamma_{\rm pia}}$, on the skull by $\mathcal{F}_{\Gamma_{\rm skull}}$, and on the spinal canal by $\mathcal{F}_{\Gamma_{\rm SAS}}$. Note that $\mathcal{F}_i \cap \mathcal{F}_{\Gamma_{\rm pia}} = \emptyset $. The union of all the facets is denoted by $\mathcal{F}_h$. The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is denoted by $\mathcal{E}_{h}$.

@@

We follow \cite{hong2016robust} and approximate the velocity field $u$ and the pressure $p$ with the following finite element spaces. 
\begin{align}
\bm V_h  & = \{ \bm v  \in H(\mathrm{div}; \Omega_{\mathrm{CSF}}): \; \bm v \vert_{E} \in \mathbb{BDM}^2 (E),  \; E \in \mathcal{E}_h; \;  \bm v \cdot \bm n = 0 \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull^l}} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} \text{ on } \Gamma_{\mathrm{LV}} \} \\ 
S_h  & = \{q \in L^2(\Omega):  \;  q \vert_E \in \mathbb{P}^1(E),  \; E \in \mathcal{E}_h \}. 
\end{align}
Here, $H(\mathrm{div};\Omega)$ is the space of $L^2$ vector fields with $L^2$ divergence,  $\mathbb{BDM}^2$ is the Brezzi-Douglas--Marini element \cite{brezzi1987mixed} of order 2, and  $\bm n$ (resp. $\bm \tau$) is the unit outward normal (resp. tangent) vector.  
Given any vector $\bm v$, the normal and tangential components on each facet  are denoted by 
\[ 
\bm v_n = (\bm v \cdot \bm n)\bm n , \quad \bm v_t = (\bm v - \bm v_n). 
\]
Since $V_h \subset H(\mathrm{div}, \Omega_{\rm{CSF}})$, then $[\bm v_n] = 0 $ on $\mathcal{F}_{h,c}$. Continuity in the tangential component is enforced weakly via interior penalization. For notational simplicity, denote by $\mathcal{F}_{\Gamma}$ the facets lying on $\Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull}} \cup \Gamma_{\rm{SSAS}} \cup \Gamma_{\mathrm{LV}}$. Define the form 
\begin{multline}
\mathcal{A}_h (\bm u, \bm v) = \sum_{E \in \mathcal{E}_h} \int_{E} \mu \nabla u : \nabla v - \sum_{
F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma} 
} \int_{F} \mu \{\nabla u\}\cdot \bm n_F \cdot [\bm v_t] 
\\ - \sum_{
F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma} 
} \int_{F} \mu \{\nabla v\}\cdot \bm n_F \cdot [\bm u_t]
+ \sum_{
F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma} 
} \frac{\sigma \mu}{h} \int_{F} [\bm u_t ] \cdot [\bm v_t].   
\end{multline}
We set $\sigma =20$. The discretization for the CSF flow, is then to find $(\bm u_h, p_h) \in \bm V_h \times S_h$ such  that 
\begin{align}
\mathcal{A}_h(u_h, v_h)  + \sum_{F \in \mathcal{F}_{\Gamma_{\mathrm{skull}}}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n) - \int_{\Omega} \nabla \cdot \bm v_h \, p_h  & = 0, \quad \forall \bm v_h \in \bm V_h   \\ 
(\nabla \cdot \bm u_h , q_h ) & = 0, \quad \forall q_h \in S_h.
\end{align}

\subsubsection{PVS flow}

@@

The weak form reads: Find $(\hat{u}, \hat{p}) \in V_h^{\rm PVS} \times Q_h^{\rm PVS}$ such that

\begin{align}
(A \hat{u}, v) + \left(v, \frac{A \kappa}{\mu} \partial_s \hat{p}  \right) &= 0 \quad \forall v \in V_h^{\rm PVS} \\
(A \hat{u}, \partial_s q ) &= 0 \quad  \forall q \in  Q_h^{\rm PVS}
\end{align}

@@

\subsection{Modelling dispersion in the CSF spaces}
\label{sec:supp:dispersion}
